### 2017-04-10 paper/oopsla17-submission/evaluation-accuracy.tex
Wes pass over the paper. Looking good from here, will send email to Eric.

Comparing source locations from multiple tools with their own parsers is
not trivial.
%
To ensure a fair comparison when evaluating \mycroft and
\sherrloc, we removed from the dataset programs where they predicted
locations that our oracle could not match with a program expression ---
6--8\% of programs for \mycroft and 3--4\% for \sherrloc.
%
We also do not consider programs where \mycroft or \sherrloc timed out
(after one minute) or where they encountered an unsupported language
feature --- another 5\% for \mycroft and 12--13\% for \sherrloc. This
experimental design gives the state of the art tools the ``benefit of the
doubt''.

>>>>

Comparing source locations from multiple tools with their own parsers is
not trivial.
%
To ensure a fair comparison when evaluating \mycroft and
\sherrloc, we removed from the dataset programs where they predicted
locations that our oracle could not match with a program expression:
6--8\% of programs for \mycroft and 3--4\% for \sherrloc.
%
We also do not consider programs where \mycroft or \sherrloc timed out
(after one minute) or where they encountered an unsupported language
feature: another 5\% for \mycroft and 12--13\% for \sherrloc. This
experimental design gives the state of the art tools the ``benefit of the
doubt''.

### 2017-04-10 paper/oopsla17-submission/evaluation.tex
Wes pass over the paper. Looking good from here, will send email to Eric.

\paragraph{Feature Selection}
We extract a set of 277 features from each sub-expression in a student
program, including:
%
\begin{enumerate}
\item 44 local syntactic features. In addition to the syntax of \lang,
  we support the full range of arithmetic operators (integer and
  floating point), equality and comparison operators, character and
  string literals, and a user-defined |expr| type of simple arithmetic
  expressions. We discuss the challenge of supporting other
  user-defined types in \autoref{sec:discussion}.
\item 176 contextual syntactic features. For each sub-expression we
  additionally extract the local syntactic features of its parent and
  first, second, and third (left-to-right) children. If an expression
  does not have a parent or children, these features will simply be
  disabled. If an expression has more than three children, the
  classifiers will receive no information about the additional
  children.
\item 55 typing features. In addition to the types of \lang, we support
  |int|s, |float|s, |char|s, |string|s, and the user-defined |expr|
  mentioned above. These features are extracted for each sub-expression
  and for the contextual sub-expressions.
\item One feature denoting the size of each sub-expression.
\item One feature denoting whether each sub-expression is part of the
  minimal type error slice. We use this feature as a ``hard''
  constraint, sub-expressions that are not part of the minimal slice
  will be preemptively discarded, we justify this decision in
  \autoref{sec:feature-utility}.
\end{enumerate}

@@

>>>>

\paragraph{Feature Selection}
We extract a set of 277 features from each sub-expression in a student
program, including:
%
\begin{enumerate}
\item 44 local syntactic features. In addition to the syntax of \lang,
  we support the full range of arithmetic operators (integer and
  floating point), equality and comparison operators, character and
  string literals, and a user-defined |expr| type of simple arithmetic
  expressions. We discuss the challenge of supporting other
  user-defined types in \autoref{sec:discussion}.
\item 176 contextual syntactic features. For each sub-expression we
  additionally extract the local syntactic features of its parent and
  first, second, and third (left-to-right) children. If an expression
  does not have a parent or children, these features will simply be
  disabled. If an expression has more than three children, the
  classifiers will receive no information about the additional
  children.
\item 55 typing features. In addition to the types of \lang, we support
  |int|s, |float|s, |char|s, |string|s, and the user-defined |expr|
  mentioned above. These features are extracted for each sub-expression
  and for the contextual sub-expressions.
\item One feature denoting the size of each sub-expression.
\item One feature denoting whether each sub-expression is part of the
  minimal type error slice. We use this feature as a ``hard''
  constraint, sub-expressions that are not part of the minimal slice
  will be preemptively discarded. We justify this decision in
  \autoref{sec:feature-utility}.
\end{enumerate}

@@

A related threat to construct validity is our definition of the immedate
next well-typed program as the intended ground truth answer (see
\autoref{sec:overview}, Challenge 2). Students may, in theory, submit
intermediate well-typed program ``rewrites'' between the original ill-typed
program and the final intended answer. Our approach to discarding outliers
(see \autoref{sec:evaluation}) is designed to mitigate this threat.

### 2017-04-10 paper/oopsla17-submission/introduction.tex
Wes pass over the paper. Looking good from here, will send email to Eric.

\paragraph{\textbf{(1) A Dataset of Novice \ocaml Programs}}
Our first contribution is a large dataset of novice \ocaml programs,
including, but not limited to, \emph{ill-typed} programs.
%
We used an instrumented version of the \ocaml compiler to collect
student interactions over two instances of an undergraduate Programming
Languages course at
%
\begin{anonsuppress}
UC San Diego (IRB \#140608).
\end{anonsuppress}
\begin{noanonsuppress}
AUTHOR's INSTITUTION (IRB HIDDEN).
\end{noanonsuppress}
%
The data forms a time-series of programs submitted to the \ocaml
compiler, from which we extracted a set of \emph{ill-typed} programs
and their \emph{subsequent fixes}.
%
Our dataset comprises over 4,500 such pairs of programs, and many more
programs that are either well-typed or lack a corresponding fix.
%
The entire dataset (both the raw time-series and our extracted pairs of
ill-typed programs and fixes) is available at
%
\begin{anonsuppress}
\ES{TODO: finally post the data somehwere..}.
\end{anonsuppress}
\begin{noanonsuppress}
HIDDEN FOR DOUBLE-BLIND REVIEW.
\end{noanonsuppress}
%

>>>>

\paragraph{\textbf{(1) A Dataset of Novice \ocaml Programs}}
Our first contribution is a large dataset of novice \ocaml programs,
including, but not limited to, \emph{ill-typed} programs.
%
We used an instrumented version of the \ocaml compiler to collect
student interactions over two instances of an undergraduate Programming
Languages course at
%
\begin{anonsuppress}
UC San Diego (IRB \#140608).
\end{anonsuppress}
\begin{noanonsuppress}
AUTHOR's INSTITUTION (IRB HIDDEN).
\end{noanonsuppress}
%
The data consist of a time-series of programs submitted to the \ocaml
compiler, from which we extracted a set of \emph{ill-typed} programs
and their \emph{subsequent fixes}.
%
Our dataset comprises over 4,500 such pairs of programs, and many more
programs that are either well-typed or lack a corresponding fix.
%
The entire dataset (both the raw time-series and our extracted pairs of
ill-typed programs and fixes) is available at
%
\begin{anonsuppress}
\ES{TODO: finally post the data somehwere..}.
\end{anonsuppress}
\begin{noanonsuppress}
HIDDEN FOR DOUBLE-BLIND REVIEW.
\end{noanonsuppress}
%

### 2017-04-10 paper/oopsla17-submission/learning.tex
Wes pass over the paper. Looking good from here, will send email to Eric.

There are many learning algorithms to choose from, existing
on a spectrum that balances expressiveness with ease of training (and of
interpreting the learned model).
%
In this section we present four standard learning algorithms that we
will investigate:
%
(1) logistic regression,
(2) decision trees,
(3) random forests, and
(4) neural networks.
%
A thorough introduction to these techniques can be found in introductory
machine learning textbooks (\eg \ES{TODO: huma/kamalika?}).
%
We will briefly introduce each technique by describing the model it
learns, the training procedure, any hyper-parameters that must be tuned
by a human, and a summary of its advantages and disadvantages.

>>>>

There are many learning algorithms to choose from, existing
on a spectrum that balances expressiveness with ease of training (and of
interpreting the learned model).
%
In this section we consider four standard learning algorithms: 
%
(1) logistic regression,
(2) decision trees,
(3) random forests, and
(4) neural networks.
%
A thorough introduction to these techniques can be found in introductory
machine learning textbooks (\eg \ES{TODO: huma/kamalika?}).
%
We will briefly introduce each technique by describing the model it
learns, the training procedure, any hyper-parameters that must be tuned
by a human, and a summary of its advantages and disadvantages.

FIXME: Wes thinks the entire rest of this section, up to ``4. Evaluation'', 
contains much more detail than an OOPSLA audience wants on
introductory/undergraduate machine learning. (Especially since nothing
after this point is a novel contribution we make.) Perhaps we can cut some
space by focusing on the pros and cons of each technique rather than
describing the details? This is a PL paper.

### 2017-04-10 paper/oopsla17-submission/overview.tex
Wes pass over the paper. Looking good from here, will send email to Eric.

\paragraph{\textbf{Challenge (1): How to represent programs?}}
The first and prime challenge we address is how to translate programs
into a format favorable to machine learning. 
%
Programs can be represented in many ways, from the raw text to 
richly-structured abstract syntax trees and control-flow graphs, but 
representations are not directly amenable to machine learning.
%
Machine learning techniques typically take as input \emph{feature
vectors}, fixed-length vectors of real numbers where each element
represents some property of the input.
%
Thus, to apply machine learning to our problem we must
determine:
%
(1) what are interesting features of a program; and
%
(2) how can we represent a variable-size program as a fixed-length vector?

@@

\paragraph{\textbf{Challenge (2): What is the ``correct'' expression to blame?}}
As we saw in \autoref{fig:sumList}, type errors admit many different
fixes to different expressions, each resulting in a different program.
%
How is the type checker to judge which expressions are more likely to be
incorrect than others?
%
Prior work has often enlisted expert users to manually judge ill-typed
programs and determine the ``correct'' fix \citep[e.g.][]{Loncaric2016-uk}, but
this approach does not scale well to a dataset large enough to support
machine learning.
%
Furthermore, while expert users have intimate knowledge of the type system,
they cannot know in general what novice users \emph{intended} to do.

>>>>

\paragraph{\textbf{Challenge (1): How to represent programs?}}
The first and prime challenge we address is how to translate programs
into a format favorable to machine learning. 
%
Programs can be represented in various ways, from the raw text to 
richly-structured abstract syntax trees and control-flow graphs, but 
many representations are not directly amenable to machine learning.
%
Machine learning techniques typically take as input \emph{feature
vectors}, fixed-length vectors of real numbers where each element
represents some property of the input.
%
Thus, to apply machine learning to our problem we must
determine:
%
(1) what are interesting features of a program; and
%
(2) how can we represent a variable-size program as a fixed-length vector?

@@

\paragraph{\textbf{Challenge (2): What is the ``correct'' expression to blame?}}
As we saw in \autoref{fig:sumList}, type errors admit many different
fixes to different expressions, each resulting in a different program.
%
How is the type checker to judge which expressions are more likely to be
incorrect than others?
%
Prior work has often enlisted expert users to manually judge ill-typed
programs and determine the ``correct'' fix \citep[\eg][]{Loncaric2016-uk}, but
this approach does not scale well to a dataset large enough to support
machine learning.
%
Furthermore, while expert users have intimate knowledge of the type system,
they cannot know in general what novice users \emph{intended} to do.
